---
title: "LRMoE Fitting"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(doParallel)
library(parallel)
library(actuar)
library(rmutil)

# LRMoE package
library(devtools)
install_github("sparktseung/LRMoE")
library(LRMoE)

# French Motor Third-Party Liability datasets
library(CASdatasets)
data(freMTPLfreq, freMTPLsev)
```

# Introduction/Recap

We work with the French Motor Third-Party Liability datasets "freMTPLfreq" and "freMTPLsev".

"freMTPLfreq" contains 10 columns:
1. PolicyID: The unique identifier policy ID (used to link with the claims dataset).
2. ClaimNb: Number of claims during the exposure period.
3. Exposure: The period of exposure for a policy, in years.
4. Power: The power of the car (ordered categorical).
5. CarAge: The vehicle age, in years.
6. DriverAge The driver age, in years.
7. Brand: The car brand, divided into 7 groups.
8. Gas: The car gas, Diesel or regular.
9. Region: The policy region in France (based on the 1970-2015 classiﬁcation).
10. Density: The density of inhabitants (number of inhabitants per km2) in the city the policyholder lives in.

"freMTPLsev" contains 2 columns:
1. PolicyID
2. ClaimAmount: The cost of the claim.

For simplicity, we neglect information on "Exposure" and "Density".  The removal of the former sets the exposure to a default of 1.  Justifications:
1. Since majority of the policies has exposure period within a year with 1 being the mode, we expect the effect of such removal to be moderate.
2. We will neglect both information throughout the whole project, putting all models on a level playing field.

We will also give a few definitions before-hand.
1. Claim: the request made by a policyholder to the insurer to compensate for a loss covered by the insurance.
2. Claim amount: the amount of money that the insurer must pay.
3. Claim frequency: the number of claims per policy.
4. Claim severity: the size/amount per claim.




# Task Description

This session of the project looks at claim modelling using the Logit-weighted Reduced Mixture of Experts (LRMoE) models.  LRMoE is a class of Mixture of Experts (MoE) models with the following specifications:
1. Employ a logit-type gating function.
2. Remove the regression structure in the expert functions, such that only the gating functions regress on the covariates.

Our goal is to predict the expected value (or mean) of the total claim amount per policy, i.e. the total costs of all claims incurred by each policyholder.

There are two common modelling approaches:
1. Model the claim frequency and severity separately, then multiply the predictions of both to produce the total claim amount.
2. Model the total claim amount per policy directly.

We will illustrate both approaches.

We will utilize the package "LRMoE".  Currently the LRMoE model does not support random effects yet.




# Data Processing

LRMoE requires the input data to be structured in a specific manner.  We will start with data processing.

Prepare dataframes for frequency, severity and aggregate claim amounts.

For aggregate claim amounts:
1. Aggregate the claim amounts by policies (using PolicyID).
2. Combine the two datasets by matching PolicyID.  If a specific policy has no claim, its claim amount appears as "NA" so we replace these values by 0.
3. Drop covariates "Exposure" and "Density".

For severity, combine the two datasets by matching PolicyID; only include non-zero claim amounts and corresponding policyholder information.

Note: Severity dataframe is smaller (have fewer observations) than the other two.

```{r Dataframes}
# Aggregate claim amounts
sev.aggre = aggregate(freMTPLsev$ClaimAmount, by = list(PolicyID = freMTPLsev$PolicyID), FUN = "sum")
colnames(sev.aggre)[2] = "ClaimAmount"

# Match two datasets by PolicyID. NA values correspond to no claims.
df.all = merge(freMTPLfreq, sev.aggre, by = "PolicyID", all = TRUE)
df.all$ClaimAmount[is.na(df.all$ClaimAmount)] = 0
df.all = df.all[,-which(names(df.all) %in% c("Exposure", "Density"))]

# Frequency dataframe
df.freq = freMTPLfreq[,-which(names(freMTPLfreq) %in% c("Exposure", "Density"))]

# Severity dataframe, only non-zero claim amounts and corresponding policyholder information
df.sev = merge(freMTPLsev, freMTPLfreq, by = "PolicyID", all = FALSE)
df.sev = df.sev[,-which(names(df.sev) %in% c("Exposure", "Density"))]

jpeg(file="exploratory1.jpeg", width=600, height=400)
hist(log(df.all$ClaimAmount), breaks=500,
     xlab="log(Claim Amount)", main="History of log(Claim Amount)")
dev.off()
```

We can see that aggregate claim amounts seem to be multi-modal, supporting the use of LRMoE.

```{r Y matrices building}
# Build matrix Y (claim information); no censoring nor truncation

sample.size = nrow(df.all)

Y.freq = matrix(c(rep(0, sample.size), # = tl
             df.freq$ClaimNb, # = yl
             df.freq$ClaimNb, # = yu
             rep(Inf, sample.size) # = tu
             ), ncol = 4, byrow = FALSE)

Y.sev = matrix(c(rep(0, nrow(df.sev)), # = tl
             df.sev$ClaimAmount, # = yl
             df.sev$ClaimAmount, # = yu
             rep(Inf, nrow(df.sev)) # = tu
             ), ncol = 4, byrow = FALSE)

Y = matrix(c(rep(0, sample.size), # = tl
             df.all$ClaimAmount, # = yl
             df.all$ClaimAmount, # = yu
             rep(Inf, sample.size) # = tu
             ), ncol = 4, byrow = FALSE)
```

```{r X matrices building}
# Build matrix X (covariates)

X.continuous = cbind(df.all$CarAge, df.all$DriverAge)
X.power = model.matrix(~df.all$Power, data = df.all) # Default is 'd'
X.brand = model.matrix(~df.all$Brand, data = df.all) # Default is 'Fiat'
X.gas = model.matrix(~df.all$Gas, data = df.all) # Default is 'Diesel'
X.region = model.matrix(~df.all$Region, data = df.all) # Default is 'Aquitaine'
X = matrix(cbind(rep(1, sample.size), # Intercept
                 X.continuous, X.power[,-1], X.brand[,-1], X.gas[,-1], X.region[,-1]), 
           nrow = sample.size, byrow = FALSE)

colnames(X) = c("Intercept", "CarAge", "DriverAge",
                "Powere", "Powerf", "Powerg", "Powerh", "Poweri", "Powerj", 
                "Powerk", "Powerl", "Powerm", "Powern", "Powero",
                "BrandJK", "BrandMCB", "BrandOGF", "BrandOther", "BrandRNC", "BrandVAS",
                "GasRegular",
                "RegionBN", "RegionB", "RegionC", "RegionHN", "RegionIF",
                "RegionL", "RegionNPC", "RegionPL", "RegionPC")


X.sev.continuous = cbind(df.sev$CarAge, df.sev$DriverAge)
X.sev.power = model.matrix(~df.sev$Power, data = df.sev) # Default is 'd'
X.sev.brand = model.matrix(~df.sev$Brand, data = df.sev) # Default is 'Fiat'
X.sev.gas = model.matrix(~df.sev$Gas, data = df.sev) # Default is 'Diesel'
X.sev.region = model.matrix(~df.sev$Region, data = df.sev) # Default is 'Aquitaine'
X.sev = matrix(cbind(rep(1, nrow(df.sev)), # Intercept
                 X.sev.continuous, X.sev.power[,-1], X.sev.brand[,-1], X.sev.gas[,-1], X.sev.region[,-1]), 
           nrow = nrow(df.sev), byrow = FALSE)

colnames(X.sev) = c("Intercept", "CarAge", "DriverAge",
                "Powere", "Powerf", "Powerg", "Powerh", "Poweri", "Powerj", 
                "Powerk", "Powerl", "Powerm", "Powern", "Powero",
                "BrandJK", "BrandMCB", "BrandOGF", "BrandOther", "BrandRNC", "BrandVAS",
                "GasRegular",
                "RegionBN", "RegionB", "RegionC", "RegionHN", "RegionIF",
                "RegionL", "RegionNPC", "RegionPL", "RegionPC")
```

Intercept is 1.

Power: default is "d".

Brand: default is "Fiat"; JK is "Japanese (except Nissan) or Korean"; MCB is "Mercedes, Chrysler or BMW"; OGF is "Opel, General Motors or Ford"; Other is "other"; RNC is "Renault, Nissan or Citroen"; VAS is "Volkswagen, Audi, Skoda or Seat".

GasRegular: default is "Diesel", the other is "Regular".

Region: default is "Aquitaine"; BN is "Basse-Normandie"; B is "Bretagne"; C is "Centre"; HN is "Haute-Normandie"; IF is "Ile-de-France"; L is "Limousin"; NPC is "Nord-Pas-de-Calais"; PL is "Pays-de-la-Loire"; PC is "Poitou-Charentes".

```{r Save and load matrices}
# save(X, file = "X.Rda")
# save(X.sev, file = "X.sev.Rda")
# save(Y, file = "Y.Rda")
# save(Y.freq, file = "Y.freq.Rda")
# save(Y.sev, file = "Y.sev.Rda")

load("X.Rda")
load("X.sev.Rda")
load("Y.Rda")
load("Y.sev.Rda")
load("Y.freq.Rda")
```




# Parameter Initialization

```{r Normalize dataframes}
# Normalize data for k-means clustering

df.all.norm = df.all[,-which(names(df.all) %in% c("PolicyID", "ClaimNb"))]
df.all.norm$CarAge = (df.all$CarAge - mean(df.all$CarAge))/sd(df.all$CarAge)
df.all.norm$DriverAge = (df.all$DriverAge - mean(df.all$DriverAge))/sd(df.all$DriverAge)
df.all.norm$ClaimAmount = (df.all$ClaimAmount - mean(df.all$ClaimAmount))/sd(df.all$ClaimAmount)

df.freq.norm = df.freq[,-which(names(df.freq) %in% c("PolicyID", "ClaimAmount"))]
df.freq.norm$CarAge = (df.freq$CarAge - mean(df.freq$CarAge))/sd(df.freq$CarAge)
df.freq.norm$DriverAge = (df.freq$DriverAge - mean(df.freq$DriverAge))/sd(df.freq$DriverAge)
df.freq.norm$ClaimNb = (df.freq$ClaimNb - mean(df.freq$ClaimNb))/sd(df.freq$ClaimNb)

df.sev.norm = df.sev[,-which(names(df.sev) %in% c("PolicyID", "ClaimNb"))]
df.sev.norm$CarAge = (df.sev$CarAge - mean(df.sev$CarAge))/sd(df.sev$CarAge)
df.sev.norm$DriverAge = (df.sev$DriverAge - mean(df.sev$DriverAge))/sd(df.sev$DriverAge)
df.sev.norm$ClaimAmount = (df.sev$ClaimAmount - mean(df.sev$ClaimAmount))/sd(df.sev$ClaimAmount)
```

Before performing k-means clustering, we try to determine the optimal number of clusters using the "Elbow Method".

```{r Determine number of clusters}
set.seed(7777)
# Compute and plot wss for k = 1 to k = 10.
wss <- sapply(1:10,
              function(k){kmeans(data.matrix(df.all.norm), k)$tot.withinss})

# jpeg(file="wss.jpeg", width=600, height=400)
plot(1:10, wss,
     type="b", pch = 19, frame = FALSE, 
     xlab="Number of Clusters K",
     ylab="Total Within-Clusters Sum of Squares")
# dev.off()
```

There seems to be kinks at k=4 and at k=6 (for frequency, severity and aggregate), indicating that the optimal number of clusters may be 4 or 6.

Following this reasoning, we mainly fit models with 4-6 clusters/latent groups.

For aggregate claim amounts, we fitted models with 3-6 latent groups.

For frequency, we fitted models with 4 and 5 latent groups.  Improvement from 4 to 5 latent groups is not particularly significant, so we did not go on to fit 6 latent groups.

For severity, we fitted models with 4, 6 and 8 latent groups.  This is motivated by the large range of claim amounts.  However, the chi-square evaluation below suggest that more latent groups do not necessarily provide a better fit.

```{r Init analysis - total}
set.seed(7777)
norm.km.cluster.3 = kmeans(data.matrix(df.all.norm), 3)
norm.init.analysis.3 = cluster.mm.severity(df.all$ClaimAmount, norm.km.cluster.3$cluster)
save(norm.init.analysis.3, file = "Init3.Rda")

set.seed(7777)
norm.km.cluster.4 = kmeans(data.matrix(df.all.norm), 4)
norm.init.analysis.4 = cluster.mm.severity(df.all$ClaimAmount, norm.km.cluster.4$cluster)
save(norm.init.analysis.4, file = "Init4.Rda")

set.seed(7777)
norm.km.cluster.5 = kmeans(data.matrix(df.all.norm), 5)
norm.init.analysis.5 = cluster.mm.severity(df.all$ClaimAmount, norm.km.cluster.5$cluster)
save(norm.init.analysis.5, file = "Init5.Rda")

set.seed(7777)
norm.km.cluster.6 = kmeans(data.matrix(df.all.norm), 6)
norm.init.analysis.6 = cluster.mm.severity(df.all$ClaimAmount, norm.km.cluster.6$cluster)
save(norm.init.analysis.6, file = "Init6.Rda")
```

```{r Init analysis - frequency}
set.seed(7777)
norm.km.cluster.4.freq = kmeans(data.matrix(df.freq.norm), 4)
norm.init.analysis.4.freq = cluster.mm.frequency(df.freq$ClaimNb, norm.km.cluster.4.freq$cluster)
save(norm.init.analysis.4.freq, file = "freq.Init4.Rda")

set.seed(7777)
norm.km.cluster.5.freq = kmeans(data.matrix(df.freq.norm), 5)
norm.init.analysis.5.freq = cluster.mm.frequency(df.freq$ClaimNb, norm.km.cluster.5.freq$cluster)
save(norm.init.analysis.5.freq, file = "freq.Init5.Rda")

set.seed(7777)
norm.km.cluster.6.freq = kmeans(data.matrix(df.freq.norm), 6)
norm.init.analysis.6.freq = cluster.mm.frequency(df.freq$ClaimNb, norm.km.cluster.6.freq$cluster)
save(norm.init.analysis.6.freq, file = "freq.Init6.Rda")
```

```{r Init analysis - severity}
set.seed(7777)
norm.km.cluster.4.sev = kmeans(data.matrix(df.sev.norm), 4)
norm.init.analysis.4.sev = cluster.mm.severity(df.sev$ClaimAmount, norm.km.cluster.4.sev$cluster)
save(norm.init.analysis.4.sev, file = "sev.Init4.Rda")

set.seed(7777)
norm.km.cluster.6.sev = kmeans(data.matrix(df.sev.norm), 6)
norm.init.analysis.6.sev = cluster.mm.severity(df.sev$ClaimAmount, norm.km.cluster.6.sev$cluster)
save(norm.init.analysis.6.sev, file = "sev.Init6.Rda")

set.seed(7777)
norm.km.cluster.8.sev = kmeans(data.matrix(df.sev.norm), 8)
norm.init.analysis.8.sev = cluster.mm.severity(df.sev$ClaimAmount, norm.km.cluster.8.sev$cluster)
save(norm.init.analysis.8.sev, file = "sev.Init8.Rda")
```




# Model Fitting

By default, the fitting function in the package imposes a penalty on the magnitude of parameters to avoid spurious models with extremely large or small parameter values.

For the same reason, initialization with extremely large or small parameter values should be avoided as well.  

Below are the codes for some of the fitted models.

```{r Frequency model}
model.list = NULL

dim.m = 1
n.covar = 30

# zero-inflated Gamma Count for frequency, with 5 latent classes

n.comp = 5

model.name = "freq.ZIgammacount5"
comp.dist = matrix( c("ZI-gammacount", "ZI-gammacount", "ZI-gammacount", "ZI-gammacount", "ZI-gammacount"),
                    nrow = dim.m, byrow = TRUE)

# Assuming no knowledge of covariate influence on response,
# initialize only the intercept coefficient, where g is the reference group
alpha.init = matrix(0, nrow = n.comp, ncol = n.covar)
alpha.init[,1] = c(log(0.11), log(0.15), log(0.13), log(0.15), log(0.46)) - log(0.46)

zero.init = matrix( c(0.96, 0.97, 0.96, 0.96, 0.96),
                    nrow = dim.m, byrow = TRUE)

params.init = list( list( c(1.094624, 22.877434), c(1.175911, 12.495032), c(1.104364, 20.169784), 
                          c(1.128669, 16.464637), c(1.090487, 23.243859) ))

hyper.alpha = 5

hyper.params = list( list( c(5, 5, 5, 5),
                           c(5, 5, 5, 5),
                           c(5, 5, 5, 5),
                           c(5, 5, 5, 5),
                           c(5, 5, 5, 5)))

model.list[[length(model.list)+1]] = list(model.name = model.name,
                                          n.comp = n.comp, comp.dist = comp.dist,
                                          alpha.init = alpha.init, zero.init = zero.init,
                                          params.init = params.init,
                                          hyper.alpha = hyper.alpha, hyper.params = hyper.params)
```

```{r Severity model}
model.list = NULL

dim.m = 1
n.covar = 30

# Lognormal for severity, with 6 latent classes

n.comp = 6

model.name = "sev.lnorm6"
comp.dist = matrix( c("lnorm", "lnorm", "lnorm", "lnorm", "lnorm", "lnorm"),
                    nrow = dim.m, byrow = TRUE)

alpha.init = matrix(0, nrow = n.comp, ncol = n.covar)
alpha.init[,1] = c(log(0.14), log(0.11), log(0.13), log(0.42), log(0.14), log(0.06)) - log(0.06)

zero.init = matrix( c(0, 0, 0, 0, 0, 0),
                    nrow = dim.m, byrow = TRUE)

params.init = list( list( c(6.27, 1.59), c(5.28, 2.35), c(6.40, 1.47),
                          c(5.57, 2.08), c(6.44, 1.40), c(6.76, 1.09) ))

hyper.alpha = 5

hyper.params = list( list( c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5)))

model.list[[length(model.list)+1]] = list(model.name = model.name,
                                          n.comp = n.comp, comp.dist = comp.dist,
                                          alpha.init = alpha.init, zero.init = zero.init,
                                          params.init = params.init,
                                          hyper.alpha = hyper.alpha, hyper.params = hyper.params)

# Gamma for severity, with 6 latent classes

n.comp = 6

model.name = "sev.gamma6"
comp.dist = matrix( c("gamma", "gamma", "gamma", "gamma", "gamma", "gamma"),
                    nrow = dim.m, byrow = TRUE)

alpha.init = matrix(0, nrow = n.comp, ncol = n.covar)
alpha.init[,1] = c(log(0.14), log(0.11), log(0.13), log(0.42), log(0.14), log(0.06)) - log(0.06)

zero.init = matrix( c(0, 0, 0, 0, 0, 0),
                    nrow = dim.m, byrow = TRUE)

params.init = list( list( c(8.81e-02, 2.12e+04), c(3.98e-03, 7.83e+05), 
                          c(1.30e-01, 1.36e+04), c(1.32e-02, 1.75e+05), 
                          c(1.63e-01, 1.02e+04), c(0.43, 3633.99) ))

hyper.alpha = 5

hyper.params = list( list( c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5),
                           c(5, 5, 5, 5, 5)))

model.list[[length(model.list)+1]] = list(model.name = model.name,
                                          n.comp = n.comp, comp.dist = comp.dist,
                                          alpha.init = alpha.init, zero.init = zero.init,
                                          params.init = params.init,
                                          hyper.alpha = hyper.alpha, hyper.params = hyper.params)
```

```{r Aggregate Model}
model.list = NULL

dim.m = 1
n.covar = 30

# zero-inflated Lognormal for aggregate claim amounts, wih 4 latent classes

n.comp = 4

model.name = "llll"
comp.dist = matrix( c("ZI-lnorm", "ZI-lnorm", "ZI-lnorm", "ZI-lnorm"),
                    nrow = dim.m, byrow = TRUE)

alpha.init = matrix(0, nrow = n.comp, ncol = n.covar)
alpha.init[,1] = c(log(0.12), log(0.18), log(0.49), log(0.21)) - log(0.21)

zero.init = matrix( c(0.96, 0.97, 0.96, 0.96),
                    nrow = dim.m, byrow = TRUE)

params.init = list( list( c(5.27, 2.34), c(6.50, 1.45), c(5.66, 2.05), c(6.33, 1.52) ))

hyper.alpha = 5

hyper.params = list( list( c(5, 5, 5),
                           c(5, 5, 5),
                           c(5, 5, 5),
                           c(5, 5, 5) ))

model.list[[length(model.list)+1]] = list(model.name = model.name,
                                          n.comp = n.comp, comp.dist = comp.dist,
                                          alpha.init = alpha.init, zero.init = zero.init,
                                          params.init = params.init,
                                          hyper.alpha = hyper.alpha, hyper.params = hyper.params)
```

```{r Function: do_fitting}
do_fitting = function(Y, X, b, model){

  # Output file names: may be modified

  model.name = toString(paste(model$model.name, sep=""))
  rda.name = toString(paste(model.name, ".Rda", sep=""))
  output.name = toString(paste(model.name, ".txt", sep=""))

  # Open new file to save intermediate update of parameter values
  sink(file = output.name, append = FALSE, type = c("output", "message"), split = FALSE)

  # Call fitting function
  tryCatch({model.fit = LRMoE::LRMoE.fit(Y = Y, X = X, n.comp = model$n.comp,
                                         comp.dist = model$comp.dist,
                                         alpha.init = model$alpha.init,
                                         zero.init = model$zero.init,
                                         params.init = model$params.init,
                                         penalty = TRUE,
                                         hyper.alpha = model$hyper.alpha,
                                         hyper.params = model$hyper.params,
                                         eps = 0.05, ecm.iter.max = 500,
                                         print = TRUE)
  save(model.fit, file = rda.name)
  },
  error=function(e){"Error!"; print("Error!")}
  )

  # Save intermediate update of parameter values
  sink()

  # Optional: use mailR to get running status. Code is omitted.

}
```

```{r Model fitting process}
# Specify how many models to fit in parallel
ncore = 1
n.run = length(model.list)

# Make computing clusters: standard procedure
cl = makePSOCKcluster(ncore)
registerDoParallel(cl)

# Call fitting functions in parallel
# Fit frequency model
ecm.table = foreach(b=1:n.run) %dopar% {do_fitting(Y.freq, X, b, model.list[[b]])}
# Fit severity model
ecm.table = foreach(b=1:n.run) %dopar% {do_fitting(Y.sev, X.sev, b, model.list[[b]])}
# Fit aggregate model
ecm.table = foreach(b=1:n.run) %dopar% {do_fitting(Y, X, b, model.list[[b]])}

# Stop computing clusters: standard procedure
stopCluster(cl)
```


## Chi-square

```{r Parameters for chi-square}
# calculate chi-square for a sample model llblll
load("1_llblll.Rda")

model.fit$params.fit

model.fit$zero.fit

model.fit$AIC
model.fit$BIC
model.fit$ll

meanlog_list = c(7.09655741886, 7.4533088886, NA, 7.05490876158, 6.229620024, 6.38022614790)
sdlog_list = c(0.03757386651, 0.5966779518, NA, 0.04118970572, 1.572756443, 0.06227624722)
zero_list = c(0.9476059027, 0.9192492805, 0.9978837217, 0.9208827225, 0.9580900511, 0.9920487088)

prob = predict.class.prob(X, model.fit$alpha.fit)

pmf = plnorm(pt, meanlog = meanlog_list, sdlog = sdlog_list)
pmf[3] = actuar::pburr(pt, shape1 = 1.411449971, shape2 = 1.000018992, scale = 24072.989814429)
# pmf = pgamma(pt, shape = shape_list, scale = scale_list)
# pdf = dgammacount.new(rep(pt, g), m_list, s_list)

combined = zero_list + (1-zero_list)*pmf

store = rowSums( sweep(prob, MARGIN = 2 , STATS = combined, FUN = "*", check.margin = TRUE))

sum(store)
```

This concludes the fitting procedure of LRMoE.

